---
title: "1.2 & 1.3"
output: html_notebook
---

# Open clean dataset
```{r}
data <- read.csv("Initial_Gene_Dataset.csv", header = TRUE, row.names = 1)
```


# Get all genes
```{r}
gene_list <- data$Gene.Symbol

gene_list <- gene_list[!duplicated(gene_list)]
```

```{r}
df <- read.delim("BIOGRID-ORGANISM-Homo_sapiens-4.2.192.tab.txt", header = TRUE, sep = "\t", dec = ".")
```

```{r}
clean_df <- df[df$ORGANISM_A_ID == "9606" & df$ORGANISM_B_ID == "9606", c("OFFICIAL_SYMBOL_A", "OFFICIAL_SYMBOL_B")]
```

## remove duplicates
```{r}
clean_df <- clean_df[!duplicated(clean_df),]
```

```{r}
clean_df <- clean_df[clean_df$OFFICIAL_SYMBOL_A != clean_df$OFFICIAL_SYMBOL_B, ]
```


```{r}
with_seeds <- clean_df[clean_df$OFFICIAL_SYMBOL_A %in% gene_list | clean_df$OFFICIAL_SYMBOL_B %in% gene_list,]
```

## retreive seconds
```{r}
genes_interct_seed <- c(with_seeds$OFFICIAL_SYMBOL_A, with_seeds$OFFICIAL_SYMBOL_B)
```

```{r}
genes_interct_seed <- genes_interct_seed[!duplicated(genes_interct_seed)]
```

```{r}
seconds <- setdiff(genes_interct_seed, gene_list)
```
## get interaction between seconds
```{r}
with_seconds <- clean_df[clean_df$OFFICIAL_SYMBOL_A %in% seconds & clean_df$OFFICIAL_SYMBOL_B %in% seconds,]
```

```{r}
all_interect <- rbind(with_seeds, with_seconds)
```

# Save
```{r}
write.csv(with_seeds, "with_seeds")
write.csv(with_seconds, "with_seconds")
write.csv(all_interect, "all_interact")
```


# 1_3
```{r}
seed_seed <- with_seeds[with_seeds$OFFICIAL_SYMBOL_A %in% gene_list & with_seeds$OFFICIAL_SYMBOL_B %in% gene_list,  ]
```

```{r}
uniprot_A <- c(rep(NA, nrow(seed_seed)))
for(i in 1:nrow(seed_seed)){
  pos = which(data$Gene.Symbol == seed_seed$OFFICIAL_SYMBOL_A[i])
  uniprot_A[i] <- data$Uniprot.AC[pos]
}

uniprot_B <- c(rep(NA, nrow(seed_seed)))
for(i in 1:nrow(seed_seed)){
  pos = which(data$Gene.Symbol == seed_seed$OFFICIAL_SYMBOL_B[i])
  uniprot_B[i] <- data$Uniprot.AC[pos]
}
```

```{r}
seed_seed$Uniprot_A <- uniprot_A 
seed_seed$Uniprot_B <- uniprot_B 

colnames(seed_seed) <- c("interactor A gene symbol", "interactor B gene symbol", "interactor A Uniprot AC", "interactor B Uniprot AC")

```


# all proteins interacting with at least one seed gene, same format as above
```{r}
seconds_with_seeds <- with_seeds[!(with_seeds$OFFICIAL_SYMBOL_A %in% gene_list & with_seeds$OFFICIAL_SYMBOL_B %in% gene_list),]
```

```{r}
library(biomaRt)
mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")
```

```{r}
uniA <- getBM(attributes=c('hgnc_symbol', "uniprot_gn_id"), 
      filters = 'hgnc_symbol', 
      values = unique(seconds_with_seeds$OFFICIAL_SYMBOL_A), 
      bmHeader = T,
      mart = mart)

```

#for each gene we have many AC possible, so we choose the first
```{r}
uniA = uniA[!duplicated(uniA$`HGNC symbol`),]
colnames(uniA)
colnames(uniA) <- c("OFFICIAL_SYMBOL_A", "Interactor A Uniprot AC")
```

#the corresponding uniprot AC are returned in alphabetical order, so we sort our dataframe to better check that everything is working fine

```{r}
seconds_with_seeds <- seconds_with_seeds[order(seconds_with_seeds$OFFICIAL_SYMBOL_A),]
```
#Now we join the to tables to attach the AC information 

```{r}
disease_interactome1 <- merge(x = seconds_with_seeds, y = uniA, by = "OFFICIAL_SYMBOL_A", all=TRUE) #outer join to avoid losing values
```
#Please note that for not all genes a Uniprot AC was found, so the final table will have some Nans (46 are missing)

```{r}
length(unique(seconds_with_seeds$OFFICIAL_SYMBOL_A))
dim(uniA)
any(is.na(disease_interactome1$'Interactor A Uniprot AC'))
```

#Now we do the same for the B column
```{r}
uniB <- getBM(attributes=c('hgnc_symbol', "uniprot_gn_id"), 
      filters = 'hgnc_symbol', 
      values = unique(seconds_with_seeds$OFFICIAL_SYMBOL_B), 
      bmHeader = T,
      mart = mart)

uniB = uniB[!duplicated(uniB$`HGNC symbol`),]
colnames(uniB) <- c("OFFICIAL_SYMBOL_B", "Interactor B Uniprot AC")
```

```{r}
seconds_with_seeds <- seconds_with_seeds[order(seconds_with_seeds$OFFICIAL_SYMBOL_B),]
```

```{r}
disease_interactome <- merge(x = disease_interactome1, y = uniB, by = "OFFICIAL_SYMBOL_B", all=TRUE) #outer join to avoid losing values
```

# Here we have the same issues but for more values
```{r}
length(unique(seconds_with_seeds$OFFICIAL_SYMBOL_B))
dim(uniB)
any(is.na(disease_interactome$'Interactor B Uniprot AC'))
```
# Let's reorder the columns and reorder the rows
```{r}
disease_interactome <- disease_interactome[,c(2,1,3,4)]
```

```{r}
disease_interactome <- disease_interactome[with(disease_interactome, order(OFFICIAL_SYMBOL_A, OFFICIAL_SYMBOL_B)), ]
```

```{r}
write.csv(disease_interactome, "disease_interactome.csv")
```


